---
title: "Homework4"
author: "Lauren Connolly"
date: "2025-02-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Step 1
```{r}
x <- 1.1
a <- 2.2
b <- 3.3
z <- c(x^a^b, (x^a)^b, 3*x^3 + 2*x^2 + 1)
print(z)
```
The values 1.1, 2.2, and 3.3 were assigned to variables x, a, and b respectively. The values of the three expressions were then assigned to z using the `c()` function.


## Step 2
```{r}
vec_2a <- c(seq(1:8), seq(from=7, to=1))
print(vec_2a)

vec_2b <- rep(1:5, times=1:5)
print(vec_2b)

vec_2c <- rep(1:5, times=5:1)
print(vec_2c)
```
Vectors from steps 2a, 2b, and 2c, are printed. Vector 2a combines a sequence from 1 to 8 with one from 7 to 1. Vector 2b repeats numbers 1 to 5 from 1 time to 5 times. Vector 2c switches the `times=` from 1:5 to 5:1 to create the opposite pattern of repetition


## Step 3 
```{r}
xy_coordinates <- runif(2)
x_cor <- xy_coordinates[1]
y_cor <- xy_coordinates[2]

r_cor <- sqrt(x_cor^2 + y_cor^2)
print(r_cor)

theta_cor <- atan(y_cor/x_cor) 
print(theta_cor)
```
The `runif(2)` creates 2 uniform numbers in a vector, and x_cor and y_cor are assigned the two different values from the vector. The r and theta coordinates are then calculated using the `sqrt()` function to calculate r from the square root and the `atan()` function to calculate theta from the arctangent.


## Step 4
```{r}
queue <- c("sheep", "fox", "owl", "ant")
print(queue) #starting queue

queue <- c(queue, "serpent")
print(queue) #a

queue <- queue[2:5]
print(queue) #b

queue <- c("donkey", queue)
print(queue) #c

queue <- queue[1:4]
print(queue) #d

queue <- c(queue[1:2], queue[4])
print(queue) #e

queue <- c(queue[1:2], "aphid", queue[3])
print(queue) #f

which(queue=="aphid") #g
```
The `c()` function is used throughout the steps to combine the queue or parts of the queue (using brackets and the numbering of the vector elements) with new elements or other parts of the queue. The new queue is printed after each step. At the end, the `which()` function is used to ask what position the aphid has in the vector. 


## Step 5
```{r}
s <- seq(1:100)
s2 <- which(s%%2 != 0 & s%%3 != 0  & s%%7 != 0)
print(s2)
```
The variable s is a sequence of all the integers between 1 and 100, and s2 finds which of those values do not have a remainder of 0 when divided by 2, 3, and 7.
