---
title: "Homework6"
author: "Lauren Connolly"
date: "2025-02-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Building Dataset
```{r}
conc1 <- rnorm(20, 1.1, 0.3)
conc5 <- rnorm(20, 1.8, 0.5)

conc_df <- data.frame(concentration = rep(c("1%", "5%"), each=20), rate = c(conc1, conc5))
print(conc_df)
```
Data was created to reflect the differences in the rates of vacuole formation between **Tetrahymena** in 1% ink versus those in 5% ink. Two variables, conc1 and conc5, were assigned vectors of normally distributed values using `rnorm()`. The sample sizes were both set to be 20, but the means and standard deviations differed to reflect how the two groups differ in reality. A dataframe was made with two columns, the rate and the concentration at which the rate occurred.

## Initial Analysis
```{r}
conc_ANOVA <- aov(rate ~ concentration, data=conc_df)
p_value <- summary(conc_ANOVA)[[1]][[1,"Pr(>F)"]]
print (p_value)
```
An ANOVA test was done on the data using `aov()`; this tested the difference in the rates bt concentration. The p-value was found in the ANOVA summary and it was printed.

## Manipulating Data
### Sample Size
```{r}
ss <- c(5, 10, 20, 40, 80)
for (i in 1:length(ss)){
  conc1 <- rnorm(ss[i], 1.1, 0.3)
  conc5 <- rnorm(ss[i], 1.8, 0.5)
  conc_df <- data.frame(concentration = rep(c("1%", "5%"), each=20), rate = c(conc1, conc5))
  conc_ANOVA <- aov(rate ~ concentration, data=conc_df)
  p_value <- summary(conc_ANOVA)[[1]][[1,"Pr(>F)"]]
  print(p_value)
}
```
To test the effect of the sample size on the p-value, the variable `ss` was first assigned a vector of 5 different sample sizes. The for loop then made data and ran an ANOVA test for all of the listed sample sizes, and it printed each p-value in order of how the sample sizes were listed.

### Mean
```{r}
m <- c(1, 2, 3, 4)
for (i in 1:length(m)){
  conc1 <- rnorm(20, 1.1+m[i], 0.3)
  conc5 <- rnorm(20, 1.8+m[i], 0.5)
  conc_df <- data.frame(concentration = rep(c("1%", "5%"), each=20), rate = c(conc1, conc5))
  conc_ANOVA <- aov(rate ~ concentration, data=conc_df)
  p_value <- summary(conc_ANOVA)[[1]][[1,"Pr(>F)"]]
  print(p_value)
}
```
To test the effect of the mean on the p-value, the variable `m` was first assigned a vector of 4 different increasing values. The for loop individually added each value to the original means for the two groups; it then made data, ran an ANOVA test for each of the new datasets, and printed each p-value in order of how the mean additions were listed in the variable `m`.

### Standard Deviation
```{r}
sd <- c(0.1, 0.2, 0.3, 0.4)
for (i in 1:length(m)){
  conc1 <- rnorm(20, 1.1, 0.3+sd[i])
  conc5 <- rnorm(20, 1.8, 0.5+sd[i])
  conc_df <- data.frame(concentration = rep(c("1%", "5%"), each=20), rate = c(conc1, conc5))
  conc_ANOVA <- aov(rate ~ concentration, data=conc_df)
  p_value <- summary(conc_ANOVA)[[1]][[1,"Pr(>F)"]]
  print(p_value)
}
```
To test the effect of the standard deviation on the p-value, the variable `sd` was first assigned a vector of 4 different increasing values. The for loop individually added each value to the original standard deviations for the two groups; it then made data, ran an ANOVA test for each of the new datasets, and printed each p-value in order of how the mean additions were listed in the variable `sd`.